---
title: "Course Project 1 Activity Monitoring"
author: "Oliver Haubner"
date: "28 Januar 2018"
output: md_document
---

***
#Loading and Preprocessing the Data
```{r, echo = TRUE, message = FALSE}
setwd("~/COURSERA/Reproducible Research/Week2/repdata%2Fdata%2Factivity")
activity <- read.csv("activity.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = "NA")
head(activity)
library(dplyr)
activity <- mutate(activity, date = as.Date(date, format = "%Y-%m-%d"))
activityTidy <- activity[!is.na(activity$steps),]
```


***
#What is the mean total number of steps taken each day?
1. Calculate the total number of steps per day
```{r, echo = TRUE, message = FALSE}
sumStep = aggregate(steps~date, activityTidy, sum)
```

2. Make a histogram of the total number of steps taken each day
```{r}
png(filename = "Plot1.png")
hist(sumStep$steps, main = "Histogram of steps per day", xlab = "Sum of steps per day")
dev.off()
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = TRUE, message = FALSE}
mean(sumStep$steps, na.rm = T)
median(sumStep$steps, na.rm = T)
```

####The daily median is **`r median(sumStep$steps, na.rm = T)`**
####The daily mean is **10766.19**

***
#What is the average daily activity pattern?

1. Make a time series plot

```{r}
avgStep = aggregate(steps~interval, activityTidy, mean)
png(filename = "Plot2.png")
plot(avgStep, type = "l", main = "Plot of average steps per interval", sub = "averaged across all observed days, ignoring NAs", xlab = "5min Intervals", ylab = "Number of steps")
dev.off()
```

2. Find interval with maximum number of steps
```{r}
maxIdx = which.max(avgStep$steps)
maxInterval = avgStep[maxIdx, 'interval']
maxAvgStep = avgStep[maxIdx, 'steps']
```
####The interval **`r maxInterval`** is the maximum interval with **`r maxAvgStep`** steps.


***
#Imputing missing values

1. Calculate and report number of missing values
```{r}
totalNA = sum(is.na(activity$steps))
```
####The total number of missing values for steps is **`r totalNA`**

2. Devise a strategy for filling out missing values  
--> Replace every NA with the calculated average for the respective time interval

3. Create new dataset with the missing data filled in
```{r}
naIdx = which(is.na(activity$steps))
naInterval = activity[naIdx, 3]
fillSteps = sapply(naInterval, function(x) {avgStep[(avgStep$interval==x), 2]})

activityFill = activity
activityFill[naIdx, 'steps'] = fillSteps

sumStepFill = aggregate(steps~date, activityFill, sum)
```

4. Histogram of new dataset (sumStepFill)
```{r}
png(filename = "Plot3.png")
hist(sumStepFill$steps, main = "Histogram of steps per day (After Imputation)", xlab = "Sum of steps per day")
dev.off()

mean(sumStepFill$steps, na.rm = T)
median(sumStepFill$steps, na.rm = T)
```


#####The mean has not changed after imputing the NAs. The median, however, has changed its value after the modification of the data.

***
#Differences in activity patterns between weekdays and weekends

```{r}
##Add new column to activityFill dataframe to indicate weekday (output is in German in my case)
activityFill2 <- activityFill
activityFill2$weekday <- weekdays(activityFill2$date)
activityFill2$weekend <- activityFill2$weekday%in%c("Samstag", "Sonntag")
head(activityFill2)

##Plot
library(lattice)
png(filename= "Plot4.png")
xyplot(steps ~ interval | weekend, data = activityFill2, layout = c(1, 2), main = "Avg. steps per interval on weekends [above] and on weekdays [below]")
dev.off()
```


